---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, message=FALSE}
library(dplyr)
library(scales)
library(ggplot2)
act_dat <- read.csv("./activity.csv")
names(act_dat) <- c('Steps', 'Date', 'Interval')
```

## What is mean total number of steps taken per day?
```{r}
sum_steps_d <- aggregate(act_dat['Steps'],
                         list(Date = act_dat$Date), 
                         sum, na.rm=TRUE)
sum_steps_d$Date <- as.Date(sum_steps_d$Date)

g <- ggplot(sum_steps_d, aes(Date, Steps))
g <- g + geom_bar(stat="identity")
g <- g + scale_x_date()
print(g)

mean_step <- mean(sum_steps_d$Steps)
median_step <- median(sum_steps_d$Steps)
```

The mean and median step per day is `r mean_step` and `r median_step` respectively.

## What is the average daily activity pattern?

```{r}
avg_steps_i <- aggregate(act_dat['Steps'],
                         list(Interval = act_dat$Interval), 
                         mean, na.rm=TRUE)

g <- ggplot(avg_steps_i, aes(Interval, Steps))
g <- g + geom_line()
print(g)

max_steps <- avg_steps_i$Interval[which.max(avg_steps_i$Steps)]
```

The maximum average number of steps occured at `r max_steps` minute interval.

## Imputing missing values

```{r}
na_cnt <- sum(is.na(act_dat$Steps))
```

The total number of NA step value is `r na_cnt`.

```{r}
act_dat_c <- act_dat
for (i in avg_steps_i$Interval){
  act_dat_c$Steps[act_dat_c$Interval == i & is.na(act_dat_c$Steps)] <- 
    as.integer(round(avg_steps_i$Steps[avg_steps_i$Interval == i], 0))
}

sum_steps_d <- aggregate(act_dat_c['Steps'],
                         list(Date = act_dat_c$Date), 
                         sum, na.rm=TRUE)
sum_steps_d$Date <- as.Date(sum_steps_d$Date)

g <- ggplot(sum_steps_d, aes(Date, Steps))
g <- g + geom_bar(stat="identity")
g <- g + scale_x_date()
print(g)

mean_step <- mean(sum_steps_d$Steps)
median_step <- median(sum_steps_d$Steps)
```

The mean and median step per day is `r mean_step` and `r median_step` respectively.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
avg_weekend <- mutate(act_dat_c, 
                      Date = as.Date(Date), 
                      Weekend = 
                        ifelse(weekdays(Date) %in% c('Saturday', 'Sunday'), 
                           'Weekend', 'Weekday')) %>% 
  group_by(Interval, Weekend) %>% 
  summarize(Steps = mean(Steps))

g <- ggplot(avg_weekend, aes(Interval, Steps))
g <- g + facet_grid(Weekend ~ ., scales='free_y') 
g <- g + geom_line()
print(g)

```

